Towards Explainability of Machine Learning Models in Insurance Pricing

### 2020-03-02 manuscript/manuscript.Rmd

Arguably, GLMs predate minimum bias procedures by a significant margin. The term 
was coined by @nelder_wedderburn_1972, but generalizations of least squares 
linear regression date back at least to the 1930s. Like minimum bias methods, 
GLMs did not become mainstream in actuarial science for some time. For example, 
the syllabus of basic education of the Casualty Actuarial Society (CAS) does not 
seem to include any mention of GLMs 
prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
seem to have received only passing mention until 2006 with the introduction of 
@anderson_2005 to the syllabus.

@@

The focus of this paper is on steps 5 and 6. In particular, rate regulators are 
concerned with whether rates are inadequate, excessive, or unfairly 
discriminatory. In many states, rate filings that exceed certain thresholds for
magnitude of rate changes or filings that make use of new or sophisticated 
predictive models may be subject to particular regulatory scrutiny. In these cases,
it is necessary to be able to explain the results of the modeling process in a 
way that is understandable without sacrificing statistical rigor.

It should be noted that communicating results is not simply a method of passing 
regulatory muster. Generating interpretable modeling output is an important - 
even essential - facet of model checking. Therefore, the techniques discussed 
in this paper may be viewed from the lens of providing useful information to 
regulators, but they should also be considered as part of a thorough vetting of 
any rating model.

@@

Of course, these comments only apply within the actuarial profession. Outside of 
the actuarial profession, communication of results may be more challenging. A 
2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
the National Association of Insurance Commissioners (NAIC) found that the plurality of
responding regulators identified "Filing complexity and/or a lack of resources 
or expertise" as a key challenge that impedes their ability to review GLMs or 
other predictive models [@naic_summer_2017]. Given that machine learning 
algorithms are generally regarded as more complex than GLMs, this implies that
the challenge of communicating machine learning model results is significant.

@@

These items are by no means exhaustive, but they pertain to the concept of model
interpretability for ratemaking that we develop next.

@@

Along the other dimension, we categorize model interpretations as global, or
model-level, and local, or instance-level. The former class provides insights
with respect to the model as a whole. Some examples of these eplanations include
variable importances and sensitivities, on average, of the predicted response
with respect to individual predictors. In our case study, questions 1 and 2 are
associated with this category. On the other hand, question 3 pertains to an
individual prediction, which would fall in the local, or instance-level,
category. In addition to individual variable attribution, we can also inquire
about what would happen to the current predicted response if we were to perturb
specific predictor variables.

@@

As noted earlier, the ordering of variables has an impact on the contributions
calculated, especially for models that are non-additive, which could cause 
inconsistent explanations. There are several approaches to ameliorate this 
phenomenon, including selecting variables with the largest contributions first,
including interactions terms, and averaging over possible orderings. The last
of these ideas is implemented by @NIPS2017_7062 using Shapley values from 
cooperative game theory, and is referred to as Shapley additive explanations 
(SHAP). These approaches are discussed further in @biecekPMVEE and its references.

@@

* How does the model impact individual insurance consumers?
* How are the predictor variables related to expected losses?
* Are the variables unfairly discriminatory?

>>>>

Arguably, GLMs predate minimum bias procedures by a significant margin. The term 
was coined by @nelder_wedderburn_1972, but generalizations of least squares 
linear regression date back at least to the 1930s. Like minimum bias methods, 
GLMs did not become mainstream in actuarial science for some time. For example, 
the syllabus of basic education of the Casualty Actuarial Society (CAS) does not 
seem to include any mention of GLMs 
prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
seem to have received only passing mention until 2006 with the introduction of 
@anderson_2005 to the syllabus. Beginning in 2016, the CAS introduced @goldburd_2016
to the syllabus, which offers a comprehensive guide to GLMs.

@@

The focus of this paper is on steps 5 and 6.
In many states, rate filings that exceed certain thresholds for
magnitude of rate changes or filings that make use of new or sophisticated 
predictive models may be subject to particular regulatory scrutiny. In these cases,
it is necessary to be able to explain the results of the modeling process in a 
way that is understandable without sacrificing statistical rigor.

It should be noted that communicating results is not simply a method of passing 
regulatory muster. Generating interpretable modeling output is an important - 
even essential - facet of model checking. Actuaries are bound by relevant standards
to be able to exercise appropriate judgment in selecting risk characteristics as
part of a risk classification system per Actuarial Standard of Practice 12
("Risk Classification"). Therefore, the techniques discussed 
in this paper may be viewed from the lens of providing useful information to 
regulators, but they should also be considered as part of a thorough vetting of 
any rating model.

Although the focus of this paper is on communication to regulators, it should be
said that selecting final rates based on indications (step 4 in the list above) may
pose a unique challenge for black-box models. This, too, provides strong motivation
for techniques that could add to the modeler's - or any stakeholder's -
understanding of the model, such as the relative importance of variables or the
shapes of response curves. Such techniques could be usefully employed in making
decisions about how best to select rates.

Similarly, although the focus of this paper is on communication in a pricing
context, the techniques explored in this paper (and many of the concerns discussed)
may also be relevant to other contexts, such as claim-level reserving / analytics
or other applications of machine learning to the insurance industry.

@@

These issues also pose particular challenges with respect to other standards.
For instance, as discussed previously, ASOP 12 requires
actuaries to be able to exercise appropriate judgment about risk classification
systems. The recent ASOP 56 ("Models") speaks to more general concerns in all
practice areas that might make use of models. ASOP 56 requires the actuary to
"make reasonable efforts to confirm that the model structure, data, assumptions,
governance and controls, and model testing and output validation are consistent
with the intended purpose." [@asop_56] All of these efforts may be hampered if
it is not possible to peer into the black box of the model.

It should also be noted that these comments only apply within the actuarial profession. Outside of 
the actuarial profession, communication of results may be more challenging. A 
2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
the National Association of Insurance Commissioners (NAIC) found that the plurality of
responding regulators identified "Filing complexity and/or a lack of resources 
or expertise" as a key challenge that impedes their ability to review GLMs or 
other predictive models [@naic_summer_2017]. Given that machine learning 
algorithms are generally regarded as more complex than GLMs, this implies that
the challenge of communicating machine learning model results is significant.

@@

The last of these items is an entire topic unto itself. The methods and
concepts introduced in this paper are useful for exploring the question of whether
rates are appropriately related to risk of loss as defined by the variables used
in the model, but there are many other aspects of discrimination-free that are
outside the scope of this paper. The methods in this paper may help in understanding
the model, which is a necessary precursor to addressing the question of unfair
discrimination.

The items in this list are by no means exhaustive, but they pertain to the concept of model
interpretability for ratemaking that we develop next.

@@

Along the other dimension, we categorize model interpretations as global, or
model-level, and local, or instance-level. The former class provides insights
with respect to the model as a whole. Some examples of these eplanations include
variable importances and sensitivities, on average, of the predicted response
with respect to individual predictors. Note that these methods may be compared to
the methods described in [@goldburd_2016], Chapter 7, which focus on global
interpretation of GLMs.

In our case study, questions 1 and 2 are
associated with global interpretations. On the other hand, question 3 pertains to an
individual prediction, which would fall in the local, or instance-level,
category. In addition to individual variable attribution, we can also inquire
about what would happen to the current predicted response if we were to perturb
specific predictor variables.

@@

As noted earlier, the ordering of variables has an impact on the contributions
calculated, especially for models that are non-additive, which could cause 
inconsistent explanations. There are several approaches to ameliorate this 
phenomenon, including selecting variables with the largest contributions first,
including interactions terms, and averaging over possible orderings. The last
of these ideas is implemented by @NIPS2017_7062 using Shapley values from 
cooperative game theory, and is referred to as Shapley additive explanations 
(SHAP). These approaches are discussed further in @biecekPMVEE and its references.
Shapley values were also used in @mango_1998, which has appeared on the CAS syllabus
starting in 2004, in the context of determining how to allocate catastrophe
risk loads between multiple accounts.

@@

* How does the model impact individual insurance consumers?
* How are the predictor variables related to expected losses?
